---
title: "How to use noaaser"
author: "Daniel Morales"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use noaaser}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(noaaser)
library(dplyr)
library(ggplot2)
filename <- system.file("extdata/earthquakes.tsv.gz", package = "noaaser")
noaase_raw <- readr::read_delim(filename, delim = "\t")
```

This package provides two types of visualization and cleaning functions to 
prepare the data for these visualizations. Exported functions are:

* `eq_clean_data`
* `eq_location_clean`
* `geom_timeline`
* `geom_timeline_label`
* `eq_map`
* `eq_create_label`

## Cleaning data

* `eq_clean_data`

Takes raw NOAA data frame and returns a clean data frame, with a date column
created by uniting the year, month, day and converting it to the Date class and 
converts LATITUDE and LONGITUDE columns to numeric class.

```r
noaase_clean <- eq_clean_data(noaase_raw)
```

* `eq_location_clean`

Takes raw NOAA data frame and returns a data frame with LOCATION_NAME column
clean, by stripping out the country name (including the colon) and converting
names to title case (as opposed to all caps).

```r
noaase_loc <- eq_location_clean(noaase_raw)
```

## Timelines

* `geom_timeline`

Timeline ggplot2 geom for NOAA significant earthquakes database.

```{r fig.width = 7, fig.align = "center"}
noaase_raw %>%
  eq_clean_data() %>%
  filter(YEAR >= 2000, YEAR <= 2016, 
         COUNTRY %in% c("ITALY", "CANADA", "GREECE")) %>%
  ggplot(ggplot2::aes(x = DATE,
                      y = COUNTRY,
                      size = as.numeric(EQ_PRIMARY),
                      color = as.numeric(TOTAL_DEATHS))) +
  geom_timeline() +
  theme_classic() +
  labs(size = "Richter scale value", color = "# deaths") +
  theme(axis.line.y = ggplot2::element_blank(),
        axis.ticks.y = ggplot2::element_blank(),
        axis.title.y = ggplot2::element_blank(),
        legend.position = "bottom") +
  ggplot2::guides(size = ggplot2::guide_legend(order = 1))
```

* `geom_timeline_label`

Provide labels for `geom_timeline`.

```{r fig.width = 7, fig.height = 4, fig.align = "center"}
noaase_raw %>%
  eq_clean_data() %>%
  eq_location_clean() %>%
  filter(YEAR >= 1900, YEAR <= 2000, COUNTRY %in% c("GREECE", "ITALY")) %>%
  ggplot(ggplot2::aes(x = DATE,
                      y = COUNTRY,
                      size = as.numeric(EQ_PRIMARY),
                      color = as.numeric(TOTAL_DEATHS),
                      label = LOCATION_NAME)) +
  geom_timeline() +
  geom_timeline_label(n_max = 1) +
  theme_classic() +
  labs(size = "Richter scale value", color = "# deaths") +
  theme(axis.line.y = ggplot2::element_blank(),
        axis.ticks.y = ggplot2::element_blank(),
        axis.title.y = ggplot2::element_blank(),
        legend.position = "bottom") +
  ggplot2::guides(size = ggplot2::guide_legend(order = 1))
```

## Maps

* `eq_map`

Maps earthquakes epicenters and annotates each point with pop-up window.

```{r fig.width = 7, fig.height = 4, fig.align = "center"}
noaase_raw %>%
  eq_clean_data() %>%
  dplyr::filter(COUNTRY == "ITALY" & lubridate::year(DATE) >= 2000) %>%
  eq_map(annot_col = "EQ_PRIMARY")
```

* `eq_create_label`

Creates an HTML label that can be used as the annotation text in the leaflet
earthquakes epicenters map from `eq_map`.

```{r fig.width = 7, fig.height = 4, fig.align = "center"}
noaase_raw %>%
  eq_clean_data() %>%
  dplyr::filter(COUNTRY == "CHILE" & lubridate::year(DATE) >= 2000) %>%
  dplyr::mutate(popup_text = eq_create_label(.)) %>%
  eq_map(annot_col = "popup_text")
```
